Memory-Efficient Mixed-Precision Implementations for\\Robust Explicit MPC

### 2019-04-10 emsoft-2019/Error_Analysis.tex
edits and comments for section 4

\section{Error Analysis}
\eva{Assuming a given precision, how to compute the error for a given controller}
The output of the controller can be affected by two main errors: (i) the controller chooses the wrong activation function because of imperfect implementation of point location algorithm, and (ii) the approximation error deriving from finite precision arithmetic used to compute the activation function itself~\cite{imperialrmpc}.\\
In an hypothetic scenario where all the measurements were done in infinite precision arithmetic and without any uncertainty, we assume the controller $i$ would be activated. Instead, because we cannot rely on infinite precision, controller $j$ is selected. The error committed is the difference between the output of the two branches $i$ and $j$.

The system has to be robust against a mistake in choosing controller $i$ instead of $j$ (or vice-versa), only when \qstatevarmath falls into the \texttt{tube} between $i$ and $j$, and not for all the points in the two regions. Moreover, we are interested in the worst case scenario where is maximized the difference between the activation functions of any two neighbor regions. Assuming that all points in region $i$ might be erroneously assigned to region $j$, without any  knowledge of the actual value of error \texttt{err}, is a too wide over-approximation of the existing instability.

Our goal is to build a feasible geometrical space around the border between two neighbor regions $i$ and $j$, where actually it is feasible that \statevarmath belongs to region $i$ but \qstatevarmath follows in region $j$ (or vice-versa).
We call this geometrical space the \texttt{tube}.

There are two main sources of error affecting the size of the \texttt{tube}: (i) the first one is caused by analog-digital conversion, happening just before the controller receives an estimation of the plant from the sensors. We call this error $\varepsilon_{A/D}$:

\begin{equation}\nonumber
\varepsilon_{A/D}=\frac{V_{cc}}{2^{p}-1}
\end{equation}

Where $V_{cc}$ is the reference voltage of the converter (e.g. typical 5V) and $p$ represents the number of bit of the processor.

The second error affecting the size of the \texttt{tube} is caused by (ii) the quantization of region bounds in the memory of the micro-controller. We call this error $\varepsilon_{Q}$ for error quantization.

While $\varepsilon_{A/D}$ is intrinsic in the capabilities of the device, $\varepsilon_{Q}$ depends on the precision used to store the boundaries. This second error can be regulated based on a trade-off among accuracy of the storage and memory save~\cite{memoryMPC}.

The total size of the \texttt{tube} is then:
\begin{equation}\label{eq:epsilontot}
\varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}
%\subsection{Finite Precision Implementation}

The effect of picking the wrong activation function are similar to instability in the switch reported in Listing \ref{lst:caseof}.

In the following, we assume $i$ and $j$ are the index of two generic regions in \statespace:
\begin{flalign}
\label{eq:maximization}
&\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}| = \\
&\max_{\forall i,j\;|\;neighbour(i,j)}|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{flalign}
where \statevarmath belongs to the \texttt{tube} between region $i$ and $j$.
Because of the linearity of the function $u_{i}-u_{j}$, and because of the convexity of the regions $i$ and $j$, it is enough to evaluate function (\ref{eq:maximization}) at the corner points of the \texttt{tube}, instead of solving a maximization problem.

We compute (\ref{eq:maximization}) for all pairs of neighbor regions $i$ and $j$. Two regions are neighbors when they share at least a border.

In formula $\exists\; m,n \;$such that:
\begin{equation}
(H_{i}\statevar-K_{j})_{m} = (H_{j}\statevar-K_{j})_{n}
\end{equation}
where $m$ and $n$ are the index of the two matching borders. We label with $border_{i,j}$ the matching border shared among the two regions.

Starting from the equation of $border_{i,j}$, we delimit the geometrical space where it makes sense to compute (\ref{eq:maximization}) with: 
\begin{equation}
\begin{aligned}
(border_{i,j} >= -\varepsilon) \land
(border_{i,j} <= \varepsilon)
\end{aligned} 
\end{equation}
where $\varepsilon$ is defined in (\ref{eq:epsilontot}). We call such geometrical space the \texttt{tube} between $i$ and $j$.

When \qstatevarmath belongs to the \texttt{tube}, it might be that $u_{i}$ is activated instead of $u_{j}$ (or vice-versa). Otherwise, when \qstatevarmath does not belong to the tube, no matter the error $\varepsilon$, the right activation function is activated, and we consider only the error deriving from the computation itself.

Since computing $u_{i}$ introduces approximation error because of finite precision arithmetic, the system has to be robust against an error that is:

\begin{equation}\label{eq:fperror}
err(u_{i})_{p}=|(\hat{F}_{p}-F)\qstatevar+(\hat{G}_{p}-G)|
\end{equation}

where $\hat{F}$ and $\hat{G}$ represent the rounded values (in p bits) for the infinite precision values $F$ and $G$.

In (7) the domain of \qstatevarmath are all the points in region $i$, but also all the values in the \texttt{tube} between region $i$ and any of the neighbors of $i$. Even if some points in the tube do not belong to region $i$, it might that $u_{i}$ is (erroneously) activated.

We compute (\ref{eq:fperror}) for all regions in \statespace\space and we take the maximum value of the error (worst case):

\begin{equation}\label{eq:maxfperror}
\max_{\forall \regionimath{i}\;in\;\statespace} err(u_{i})_{p}
\end{equation}

The disturbance \texttt{delta}  has to be an upper bound to the summation of (\ref{eq:maximization}) and (\ref{eq:maxfperror}).

In formula:
\begin{flalign}
\label{eq:delta}
&delta >= \\
&\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}| + \max_{\forall\;\regionimath{i}\;in\;\statespace} err(u_{i})_{p}\nonumber
\end{flalign}

>>>>

%!TEX root = main.tex
\section{Error Analysis}
\label{sec:Error_Analysis}

We first present our error analysis assuming given fixed-point word lengths,
before explaining our optimization algorithm which determines suitable word
lengths fully automatically in~\autoref{sec:Controller_Synthesis}.
\eva{The `p' below is the fixed/uniform word length? If so, please explain here}
\eva{Should we mention here the paper which does the LP-based analysis, i.e.
contrast our approach?}

The output of the controller is affected by two main sources errors which an error
analysis has to capture: 
\begin{itemize}
  \item[(i)] \emph{incorrect region selection}: an incorrect control action can be chosen
    due to analog-digital conversion and quantization of $H_i$ and $K_i$,

  \item[(ii)] \emph{approximate control output}: the control output is computed
    using quantized $F_i$ and $G_i$, accumulating errors.
\end{itemize}

We thus have to ensure that these errors remain below the error bound $\Delta$
used to design the robust MPC controller:
\begin{align}\label{eq:delta}
  \max_{\forall i,j. neighbour(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}| + \max_{\forall i.\mathcal{R}_{i}\in\mathcal{R}} err(u_{i})_{p} \le \Delta
\end{align}
where $i,j$ are the indexes of two generic regions $\mathcal{R}_{i}$ and
$\mathcal{R}_{j}$, \\ $neighbour(\mathcal{R}_{i},\mathcal{R}_{j})$ states that
two regions share at least a point, $u_{i}$ is the
control action for controller $i$, and $err(u_{i})_{p}$ is
the approximation error for computing $u_{i}$ using precision p.

\paragraph{Incorrect Region Selection}

The first part of the error in~\autoref{eq:delta} captures the error due to
selection of the wrong control action, i.e. with infinite-precision arithmetic
and in the absence of any uncertainty from hardware measurements we would choose
control action $u_i$, but due to uncertainties we actually choose control action $u_j$:
\begin{align} \label{eq:maximization}
  &\max_{\forall i,j.neighbour(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}|= \\
  &\max_{\forall i,j.neighbour(\mathcal{R}_{i},\mathcal{R}_{j})}|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{align}

%Because of the linearity of the function $u_{i}-u_{j}$, and because of the
%convexity of the regions $i$ and $j$, it is enough to evaluate function
%(\ref{eq:maximization}) at the corner points of the \texttt{tube}, instead of
%solving a maximization problem.

A naive computation of this error would consider all possible values of $x_k$.
This would lead to a gross over-approximation, however, since the controller
will choose a wrong region only in a relative narrow neighbourhood, or
\emph{tube}, around the boundary between two regions $R_i$ and $R_j$.
In~\autoref{eq:maximization}, we thus constrain $x_k$ to be within an appropriate tube:
\begin{equation}\label{eq:tube}
(border_{i,j} >= -\varepsilon) \land (border_{i,j} <= \varepsilon)
\end{equation}
where $border_{i,j}$ constrains $x_k$ to be on the border between two regions
$R_i$ and $R_j$ and $\varepsilon$ captures the uncertainty in computing the
correct region.

\eva{The below, I believe, is somehow defining $border_{i,j}$, but I don't understand
how. Please explicitly define $border_{i,j}$.}
In formula $\exists\; m,n \;$such that:
\begin{equation}
(H_{i}\statevar-K_{j})_{m} = (H_{j}\statevar-K_{j})_{n}
\end{equation}
where $m$ and $n$ are the indexes of the two hyperplanes. We label with $border_{i,j}$ the matching border shared among the two regions.

The width of the tube $\varepsilon$, i.e. the uncertainty in computing $x_k$ and
thus the correct region, is due to two sources of errors: analog-to-digital
conversion $\varepsilon_{A/D}$ and the quantization of region bounds
$\mathcal{R}$ in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

Analog conversion happens just before the controller receives the sensor input from the plant
and is given by:
\begin{equation*}
\varepsilon_{A/D}=\frac{V_{cc}}{2^{p}-1}
\end{equation*}
where $V_{cc}$ is the reference voltage of the converter (e.g. typically 5V) and
$p$ represents the number of bits.
\eva{which bits, who determines these?}

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries. This
second error can be tuned based on a trade-off between accuracy of the storage and
memory storage needed~\cite{memoryMPC}.
%
$\varepsilon_{Q}$ is given by:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} > (H-\hat{H}_{p})\qstatevar+(K-\hat{K}_{p})
\end{equation}
\eva{Shouldn't this equation have some absolute values in it?}
\eva{The equation above does not give any hint for how to compute it. I would prefer to
add the inequality to \autoref{eq:epsilontot} and give an equality here that explicitly
shows how to compute this. I assume there is some maximization going on here. Then also rephrase the below}
This inequality assures that the distance between any hyperplane represented in infinite precision (H and K), and its counter-part quantized in p bits ($\hat{H}_{p}$ and $\hat{K}_{p}$), is bounded by $\varepsilon_{Q}$.

\paragraph{Approximate Control Output}

Computing $u_{i}$ itself introduces imprecision because $u_i$ needs to be evaluated 
in finite precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(u_{i})_{p}=|(\hat{F}_{p}-F)\statevar+(\hat{G}_{p}-G)|
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the rounded values (in p bits) for the infinite precision values $F$ and $G$.

In (\ref{eq:fperror}) the domain for \statevarmath are all the points in region
$i$, together with all the values in the tube between region $i$ and
any of the neighbors of $i$. Even if some points in the tube do not belong to
region $i$, it might that $u_{i}$ is (erroneously) activated. 
\eva{$u_{i}$, or some $u_{j}$?}

\eva{This section so far says nothing about how we actually compute these errors (incl. with which tools). 
This should come here}
\paragraph{Implementation}
